Comparison of three similarity scores for bullet LEA matching\tnoteref{t1}\tnoteref{t2}

### 2020-01-17 index.tex
updated filfes, changes, etc.

\documentclass[doubleblind]{elsarticle}

@@

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num-names}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

@@

\begin{abstract}
Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data as a supplement to traditional comparison microscopy and 2D photography. Automatic comparison algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully,song3DTopographyMeasurements2014} or matching bullet striae \citep{aoas2, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}. One key aspect of validating these automatic comparison algorithms is to evaluate the performance of the algorithm on external tests, that is, using data which were not used to train the algorithm. Here, we present a discussion of the performance of the matching algorithm \citep{aoas2} in three studies conducted using different Ruger weapons. We consider the performance of three scoring measures: random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level. Cross correlation and random forest scores both result in perfect discrimination of same-source and different-source bullets. At the land-to-land level, discrimination for both cross correlation and random forest scores (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
\end{abstract}

@@

<<echo = FALSE, message = FALSE, warning = FALSE>>=
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  dpi = 600,
  fig.align = "center",
  out.width = "\\textwidth",
  cache = TRUE,
  fig.path = "figures/",
  echo = FALSE
)
options(knitr.table.format = "latex")

library(stringr)

library(tidyverse)
library(scales)
library(multidplyr) # install_github("hadley/multidplyr")
library(bulletxtrctr) # install_github("csafe-isu/bulletr")
library(gridExtra)
library(kableExtra)
@

@@

\begin{itemize}
\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut in the identified area; that is, an area where the striae are similar in the region above and below the crosscut. This excludes regions with extreme pitting, breakoff, tank rash, and other flaws that would interfere with similarity score calculation.
\item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
\item[(c)] Remove bullet curvature using a non-parametric smooth. 
\end{itemize}
A signature for a land engraved area is then defined as the sequence $S(x)$, $x = 1, ..., I$, where $I$ is the number of observed locations across the base of the bullet. \autoref{fig:signatures} shows a set of six signatures corresponding to the six land engraved areas of bullets 1 and 2 from barrel 1 of set 44 of the Hamby study \citep{hamby}. 
<<signatures, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44. Matching lands between the two bullets have been placed above each other.", fig.height=4, fig.width=10>>=
# signatures for one bullet
# h44cc <- readRDS("data/h44-ccdata.rda")
# br1 <- h44cc[1:12]
# dframe <- 1:12 %>% purrr::map_df(.f = function(i) {
#   d <- br1[[i]]
#   d <- d %>% group_by(y) %>% summarize(
#     value = mean(value, na.rm=TRUE)
#   )
#   d$id <- i
#   d %>% ungroup(y) %>% rename(
#     x = y
#   )
# })
#
# dframe$bullet <- rep(c("Bullet 1", "Bullet 2"), each = 6)[dframe$id]
# dframe$land <- paste("Land", (dframe$id %% 6 + 1))
#
# dframe <- dframe %>% group_by(bullet, land) %>% nest()
# names(dframe)[3] <- "ccdata"
#
# library(bulletxtrctr)
# dframe <- dframe %>% mutate(
#   grooves = ccdata %>%
#     purrr::map(.f = cc_locate_grooves, method = "middle",
#                adjust = 30, return_plot = TRUE)
# )
# dframe <- dframe %>% mutate(
#   sigs = purrr::map2(
#     .x = ccdata, .y = grooves,
#     .f = function(x, y) {
#       cc_get_signature(
#         ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
#     })
# )
# dframe$barrelland <- dframe$land
# idx <- which(dframe$bullet=="Bullet 2")
# dframe$landid <- parse_number(dframe$land)
# dframe$landid[idx] <- (dframe$landid[idx] +1) %% 6 + 1
# dframe$land <- paste("Land", dframe$landid)
# signatures <- dframe %>% select(bullet, land, sigs) %>% tidyr::unnest()

signatures <- read.csv("data/signatures.csv")
signatures %>%
  ggplot(aes(x = x / 1000)) +
#  geom_line(aes(y = raw_sig), colour = "grey70") +
  geom_line(aes(y = sig), colour = "grey30") +
  facet_grid(bullet ~ land) +
  ylim(c(-5, 5)) +
  theme_bw() +
  ylab("Surface measurement (in microns)") +
  xlab("Relative Location (in millimeters)")
@

@@

Cross correlation can also be used to assess similarity between two aligned signatures. Cross correlation varies between -1 and 1, but as the alignment is based on the maximized cross correlation value, in practice the cross correlations for aligned signatures are generally positive. Cross correlation values which are close to 1 indicate similarity between signatures, and may consequentially indicate that the two signatures come from the same source. Low cross correlation values indicate signatures which may originate from different sources.

The random forest presented in \citet{aoas2} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae. The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.

@@

<<demo-l2lscores, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
features <- read.csv("data/fsi-features.csv.gz")
features <- features %>% mutate(
  bullet1 = gsub("ullet ", "", bullet1),
  bullet2 = gsub("ullet ", "", bullet2)
)
b12 <- features %>% filter(
  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
)
# 
# p1 <- b12 %>%
#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
#   geom_tile() + ggtitle("Same barrel") +
#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
#   theme_bw() +
#   xlab("Bullet 2") + ylab("Bullet 1") +
#   theme(legend.position = "bottom") + coord_equal()
# 
bab <- features %>% filter(
  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
)
# p2 <- bab %>%
#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
#   geom_tile() + ggtitle("Different barrel") +
#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
#   theme_bw() +
#   xlab("Bullet 2") + ylab("Bullet 1") +
#   theme(legend.position = "bottom") + coord_equal()

l2lscores <- bind_rows(
  mutate(b12, type = "Same barrel"),
  mutate(bab, type = "Different barrel")
) %>% 
  mutate(type = factor(type, levels = c("Same barrel", "Different barrel"), ordered = T))

@

@@

<<b2b, echo = FALSE, fig.width = 5.4, fig.height = 3, out.width = ".6\\textwidth", fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons. On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels).">>=
# grid.arrange(p1, p2, ncol = 2, padding = 2)
ggplot(data = l2lscores, aes(x = land1, y = land2, fill = rfscore)) +
  geom_tile() + 
  scale_fill_gradient2("Random Forest\n(RF) score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
  theme_bw() +
  xlab("Bullet 2") + ylab("Bullet 1") +
  theme(legend.position = "right") + coord_equal() + facet_wrap(~type)
@

@@

<<sam-sketch, out.width='\\textwidth', fig.width=8, fig.height = 3, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands.">>=
df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
df2$case1 <- df2$x == ((df2$y + 5) %% 6) + 1
df2$case2 <- df2$x == ((df2$y + 0) %% 6) + 1
df2$case3 <- df2$x == ((df2$y + 1) %% 6) + 1
df2$case4 <- df2$x == ((df2$y + 2) %% 6) + 1
df2$case5 <- df2$x == ((df2$y + 3) %% 6) + 1
df2$case6 <- df2$x == ((df2$y + 4) %% 6) + 1

df2 %>%
  gather("case", "values", starts_with("case")) %>%
  mutate(case = gsub("case", "", case)) %>%
  mutate(case = paste0("k = ", as.numeric(case) - 1)) %>%
  ggplot(aes(x = factor(x), y = factor(y, levels = 1:6), fill = values)) +
  geom_tile(colour = "grey20") +
  facet_wrap(~case, ncol = 6) +
  xlab("Bullet 2") + ylab("Bullet 1") +
  scale_x_discrete(labels = paste0("L", 1:6)) +
  scale_y_discrete(labels = paste0("L", 1:6)) +
  scale_fill_manual("Cell included in SA(A,k)", values = c("white", "grey20")) +
  coord_equal() +
  theme(legend.position = "bottom")
@

@@

<<h44, echo=FALSE, fig.width = 8, fig.height=4, fig.cap="Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).">>=
h44features <- read.csv("data/h44-features.csv")
h44features <- h44features %>% mutate(
  lys = overlap*signature_length*1000/.645,
  cms = cms2 *lys*.645/1000
)
# tank rash lands:
#h44features <- h44features %>% mutate(
#  rfscore = replace(rfscore, (bullet1 == "I" & land1 == 6) | (bullet2 == #"I" & land2 == 6), NA)
#)
h44features <- h44features %>% mutate(
  rfscore = replace(rfscore, (barrel1 == 8 & bullet1 == "1" & land1 == 6) | (barrel2 == 8 & bullet2 == "1" & land2 == 6), NA)
)

h44nest <- h44features %>%
  filter(barrel1 == "Unk") %>%
  filter(barrel1 != barrel2 | bullet1 != bullet2) %>%
  mutate(
    barrel1 = "Unknown",
    barrel2 = as.character(barrel2),
    barrel2 = replace(barrel2, barrel2 == "Unk", "Unknown")
  ) %>%
  group_by(bullet1, barrel2, bullet2) %>%
  nest()

h44nest <- h44nest %>% mutate(
  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
    max(scores)
  }),
  sam_rf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
    max(scores)
  }),
  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
    max(scores)
  })
)
h44nest <- h44nest %>%ungroup() %>% 
  mutate(
    barrel2 = factor(barrel2, levels = c(1:10, "Unknown")),
    bullet1 = factor(bullet1, levels = rev(c(
      "K", "O", "L", "P",
      "J",
      "H", "I", "Y", "G",
      "E", "U", "X", "F",
      "T", "S"
    ))),
    bullet2 = factor(bullet2, levels = c(1:2, rev(levels(bullet1))))
  )
h44 <- h44nest %>%
  ggplot(aes(
    y = bullet1, x = bullet2,
    fill = sam_rf
  )) +
  geom_tile(size = 1) +
  facet_grid(. ~ barrel2, space = "free", scales = "free") +
  ylab("Questioned bullets") +
  xlab("Test fires from barrels 1 to 10 (left), questioned bullets (right)") +
  scale_fill_gradient2("RF score ",
    low = "darkgrey",
    high = "darkorange", midpoint = 0.35, limits = c(0, 1)
  ) +
  scale_colour_manual("same source", values = "darkorange") +

  #  scale_x_discrete(position = "top") +
  scale_y_discrete() +
  geom_tile(aes(colour = TRUE), size = .5, data = filter(h44nest, sam_rf > 0.3)) +
  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Hamby set 44") +
  theme_bw() +
  theme(legend.position = "bottom")

h44nest$samesource <- h44nest$sam_rf > 0.3

h44
@

@@

<<close, eval=FALSE, fig.width=11.25, fig.height=4.2, fig.cap="Examples of land-to-land comparisons from the Hamby-44 set. Along the top are the three bullets from different source with the highest RF scores, along the bottom are the three bullets from same source with the lowest RF score.", fig.subcap=c("Ground truth: different source", "Ground truth: same-source"), fig.align="middle", fig.ncol=1>>=
# ex2 <- h44nest %>% filter(samesource == FALSE & sam_rf >= 0.287)
# rfs2 <- ex2$sam_rf
# ex2 <- ex2 %>% mutate(
#   data = data %>% purrr::map(.f = function(d) {
#  #   browser()
#     index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
#     d$same <- ((d$land2 - d$land1) %% 6)  == index-1
#
#   #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) +
#   #                  geom_tile() +
#   #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
#   #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
#   #                      midpoint = 0.5) +
#   #     scale_colour_manual(values=c( "darkorange")) +
#   # scale_x_discrete("") +
#   # scale_y_discrete("")
#     d
#   })
# )
# ex2$example <- with(ex2, paste0(bullet1, " vs ", barrel2, "-", bullet2))
# ex2 <- unnest(ex2, data)
# ex2 %>%
#   ggplot(aes(x = land1, y = land2, fill=rfscore)) +
#   geom_tile() +
#   theme_bw() +
#   scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
#                        midpoint = 0.5) +
# #  geom_tile(colour="darkgrey", data = ex2 %>% filter(same), size=0.5) +
#   facet_grid(.~example) +
#   scale_x_discrete("") +
#   scale_y_discrete("") +
#   annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs2), hjust = 0)
#
#
#
# ex1 <- h44nest %>% filter(samesource == TRUE & sam_rf < 0.4)
# rfs <- ex1$sam_rf
# ex1 <- ex1 %>% mutate(
#   data = data %>% purrr::map(.f = function(d) {
#  #   browser()
#     index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
#     d$same <- ((d$land2 - d$land1) %% 6)  == index-1
#
#   #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) +
#   #                  geom_tile() +
#   #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
#   #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
#   #                      midpoint = 0.5) +
#   #     scale_colour_manual(values=c( "darkorange")) +
#   # scale_x_discrete("") +
#   # scale_y_discrete("")
#     d
#   })
# )
# ex1$example <- with(ex1, paste0(bullet1, " vs ", barrel2, "-", bullet2))
# ex1 <- unnest(ex1, data)
# ex1 %>%
#   ggplot(aes(x = land1, y = land2, fill=rfscore)) +
#   geom_tile() +
#   theme_bw() +
#   scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
#                        midpoint = 0.5, na.value = NA) +
#   geom_tile(colour="darkorange", data = ex1 %>% filter(same), size=0.5) +
#   facet_grid(.~example) +
#   scale_x_discrete("") +
#   scale_y_discrete("") +
#   annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs), hjust = 0)

# \autoref{fig:close} shows an example of six comparisons of bullet pairs at the land-to-land level. The panels along the top of the figure show three bullet pairs that are known to be from different barrels, the three bullet pairs along the bottom are known to be from the same barrel. The light orange frame in the bottom three panels indicate which lands correspond to one-another in the matching. It is obvious that not all of the scores are as high as we would usually see in same-source comparisons.
@

@@

<<pd, echo=FALSE, fig.width = 8, fig.height=3.56, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis). The order of the bullets on the y axis is determined by matching barrel.">>=
features <- read.csv("data/pd-features.csv.gz")
features <- features %>% mutate(
  lys = overlap*signature_length*1000/.645,
  cms = cms2 *lys*.645/1000
)
# change labels for land 3 and land 4 of bullet B3 in P7
features <- features %>% mutate(
  first = as.character(first),
  first = replace(first, first=="Gun 1-P7/B3/L4.dat", "XXX"),
  first = replace(first, first=="Gun 1-P7/B3/L3.dat", "Gun 1-P7/B3/L4.dat"),
  first = replace(first, first=="XXX", "Gun 1-P7/B3/L3.dat")
)

features <- features %>% mutate(
  KM = replace(KM, first %in% c("Gun 1-P7/B3/L3.dat", "Gun 1-P7/B3/L4.dat"), FALSE)
)
features <- features %>% mutate(
  KM = ifelse(first == "Gun 1-P7/B3/L3.dat", (second == "Unknown 1-J/L3.dat"), KM),
  KM = ifelse(first == "Gun 1-P7/B3/L4.dat", (second == "Unknown 1-J/L4.dat"), KM)
)

f2 <- features %>%
  separate(first, into = c("foo1", "foo2", "barrel1", "bullet1", "land1", "foo3"), remove = FALSE) %>%
  separate(second, into = c("foo4", "foo5", "bullet2", "land2", "foo6"), remove = FALSE) %>%
  select(-foo1, -foo2, -foo3, -foo4, -foo5, -foo6)

unknowns <- read.csv("data/pd-features-unknown.csv")
unknowns <- unknowns %>% filter(bullet1 != bullet2)
unknowns <- unknowns %>% mutate(
  bullet2 = gsub("Unknown 1-", "", bullet2),
  bullet1 = gsub("Unknown 1-", "", bullet1)
)
f3 <- rbind(
  f2 %>%
    select(
      b1, b2, barrel1, bullet1,
      land1, bullet2, land2, ccf, rfscore, cms, KM
    ) %>%
    mutate(barrel2 = "Unknown"),
  unknowns %>%
    select(
      b1, b2, bullet1,
      land1, bullet2, land2, ccf, rfscore, cms, 
    ) %>%
    mutate(barrel1 = "Unknown", barrel2 = "Unknown", KM = FALSE)
)

pdnest <- f3 %>% group_by(bullet2, barrel1, bullet1) %>% nest()

pdnest <- pdnest %>% mutate(
  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
    max(scores)
  }),
  sam_rf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
    max(scores)
  }),
  KM = data %>% purrr::map_lgl(.f = function(d) {
    any(d$KM)
  }),
  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
    max(scores)
  })
)

pdnest <- pdnest %>%ungroup() %>% 
  mutate(
    bullet2 = factor(bullet2, levels = rev(c("N", "B", "E", "T", "H", "J", "K", "Q", "Y", "Z"))),
    bullet1 = factor(bullet1, levels = c("B1", "B2", "B3", rev(levels(bullet2))))
  )
p3 <- pdnest %>%
  ggplot(aes(y = bullet2, x = bullet1, fill = sam_rf)) +
  geom_tile(size = 1) +
  facet_grid(. ~ barrel1, scales = "free", space = "free") +
  ylab("Questioned bullets") +
  xlab("Known bullets and Questioned bullets") +
  scale_fill_gradient2("RF score ",
    low = "darkgrey",
    high = "darkorange", midpoint = 0.45, limits = c(0, 1)
  ) +
  scale_colour_manual("same source", values = "darkorange") +
  scale_y_discrete() +
  theme_bw() +
  theme(legend.position = "bottom") +
  #  coord_equal() +
  geom_tile(aes(colour = TRUE),
    size = .5,
    data = filter(pdnest, KM)
  ) +
  guides(colour = guide_legend(override.aes = list(fill = NA)))

pdnest$samesource <- pdnest$KM

p3 + ggtitle("Phoenix PD set")
@

@@

<<hou, fig.width=8, fig.height = 2.67, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1>>=
features <- read.csv("data/fsi-features.csv.gz")
features <- features %>% mutate(
  lys = overlap*signature_length*1000/.645,
  cms = cms2 *lys*.645/1000
)

features <- features %>% mutate(
  bullet1 = gsub("ullet ", "", bullet1),
  bullet2 = gsub("ullet ", "", bullet2)
)

features <- features %>%
  filter(group1 == group2) %>%
  filter(barrel1 == "Unknowns") %>%
  filter(barrel1 != barrel2 | bullet1 != bullet2)

groups <- features %>%
  group_by(group1, barrel1, bullet1, barrel2, bullet2) %>%
  nest()

hounest <- groups %>% mutate(
  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
    max(scores)
  }),
  sam_rf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
    max(scores)
  }),
  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
    max(scores)
  })
)

hounest$samesource <- hounest$sam_rf > 0.3

hou1nest <- hounest %>%ungroup() %>% 
  filter(group1 == "G1") %>%
  mutate(
    bullet1 = factor(bullet1, levels = rev(c("U40", "U36", "U42", "U10", "U77", "U37", "U28", "U15"))),
    bullet2 = factor(bullet2, levels = c("B1", "B2", "B3", levels(bullet1)))
  )

hou1 <- hou1nest %>%
  ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) +
  geom_tile(size = 1) +
  facet_grid(. ~ barrel2, scales = "free", space = "free") +
  ylab("Questioned bullets") +
  xlab("Known bullets and Questioned bullets") +
  scale_fill_gradient2("RF score ",
    low = "darkgrey",
    high = "darkorange", midpoint = 0.375, limits = c(0, 1)
  ) +
  scale_y_discrete() +
  scale_colour_manual("same source", values = "darkorange") +
  geom_tile(aes(colour = TRUE), size = .5, data = filter(hou1nest, sam_rf > 0.3)) +
  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Houston set 1")

hou2nest <- hounest %>%ungroup() %>% 
  filter(group1 == "G2") %>%
  mutate(
    bullet1 = factor(bullet1, levels = rev(c("U23", "U41", "U61", "U98", "U73", "U34", "U63", "U66"))),
    bullet2 = factor(bullet2, levels = c("B1", "B2", "B3", levels(bullet1)))
  )
hou2 <- hou2nest %>%
  ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) +
  geom_tile(size = 1) +
  facet_grid(. ~ barrel2, scales = "free", space = "free") +
  ylab("Questioned bullets") +
  xlab("Known bullets and Questioned bullets") +
  scale_fill_gradient2("RF score ",
    low = "darkgrey",
    high = "darkorange", midpoint = 0.375, limits = c(0, 1)
  ) +
  scale_y_discrete() +
  scale_colour_manual("same source", values = "darkorange") +
  geom_tile(aes(colour = TRUE), size = .5, data = filter(hou2nest, sam_rf > 0.3)) +
  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Houston set 2")

hou3nest <- hounest %>%ungroup() %>% 
  filter(group1 == "G3") %>%
  mutate(
    bullet1 = factor(bullet1, levels = rev(c("U27", "U33", "U36", "U49", "U56", "U65", "U14", "U45"))),
    bullet2 = factor(bullet2, levels = c("B1", "B2", "B3", levels(bullet1)))
  )

hou3 <- hou3nest %>%
  ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) +
  geom_tile(size = 1) +
  facet_grid(. ~ barrel2, scales = "free", space = "free") +
  ylab("Questioned bullets") +
  xlab("Known bullets and Questioned bullets") +
  scale_fill_gradient2("RF score ",
    low = "darkgrey",
    high = "darkorange", midpoint = 0.375, limits = c(0, 1)
  ) +
  scale_y_discrete() +
  scale_colour_manual("same source", values = "darkorange") +
  geom_tile(aes(colour = TRUE), size = .5, data = filter(hou3nest, sam_rf > 0.3)) +
  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Houston set 3")

hou1

hou2

hou3
@

@@

<<eval=FALSE>>=
xtabs(~sameland+I(round(cms)>=6), data = all_lands)
@

>>>>

\documentclass[doubleblind]{elsarticle}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num-names}
%%%%%%%%%%%%%%%%%%%%%%%
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\begin{abstract}
Recent advances in microscopy have made it possible to collect 3D topographic data, enabling more precise virtual comparisons based on the collected 3D data as a supplement to traditional comparison microscopy and 2D photography. Automatic comparison algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully,song3DTopographyMeasurements2014} or matching bullet striae \citep{aoas2, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}. One key aspect of validating these automatic comparison algorithms is to evaluate the performance of the algorithm on external tests, that is, using data which were not used to train the algorithm. Here, we present a discussion of the performance of the matching algorithm \citep{aoas2} in three studies conducted using different Ruger weapons. We consider the performance of three scoring measures: random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level. Cross correlation and random forest scores both result in perfect discrimination of same-source and different-source bullets. At the land-to-land level, discrimination for both cross correlation and random forest scores (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
\end{abstract}

@@

@@

\begin{itemize}
\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut in the identified area; that is, an area where the striae are similar in the region above and below the crosscut. This excludes regions with extreme pitting, breakoff, tank rash, and other flaws that would interfere with similarity score calculation.
\item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
\item[(c)] Remove bullet curvature using a non-parametric smooth. 
\end{itemize}
A signature for a land engraved area is then defined as the sequence $S(x)$, $x = 1, ..., I$, where $I$ is the number of observed locations across the base of the bullet. \autoref{fig:signatures} shows a set of six signatures corresponding to the six land engraved areas of bullets 1 and 2 from barrel 1 of set 44 of the Hamby study \citep{hamby}. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/signatures-1}

}

\caption[Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44]{Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44. Matching lands between the two bullets have been placed above each other.}\label{fig:signatures}
\end{figure}

\end{knitrout}

@@

Cross correlation can also be used to assess similarity between two aligned signatures. Cross correlation varies between -1 and 1, but as the alignment is based on the maximized cross correlation value, in practice the cross correlations for aligned signatures are generally positive. Cross correlation values which are close to 1 indicate similarity between signatures, with the possibility that the two signatures come from the same source. Low cross correlation values indicate signatures which are different and thus may originate from different sources.

The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae. The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate signatures which are similar and may have originated from the same source, while scores close to 0 indicate that the aligned signatures are different and may originate from different sources.

@@

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=.6\textwidth]{figures/b2b-1}

}

\caption[Overview of land-land matching scores for two pairs of bullet-bullet comparisons]{Overview of land-land matching scores for two pairs of bullet-bullet comparisons. On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels).}\label{fig:b2b}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/sam-sketch-1}

}

\caption[Sketch of all six land-to-land sequences between two bullets with six lands]{Sketch of all six land-to-land sequences between two bullets with six lands.}\label{fig:sam-sketch}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/h44-1}

}

\caption[Hamby set 44]{Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).}\label{fig:h44}
\end{figure}

\end{knitrout}

@@

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/pd-1}

}

\caption[Phoenix study]{Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis). The order of the bullets on the y axis is determined by matching barrel.}\label{fig:pd}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\subfloat[Set 1\label{fig:hou1}]{\includegraphics[width=\textwidth]{figures/hou-1} }\newline
\subfloat[Set 2\label{fig:hou2}]{\includegraphics[width=\textwidth]{figures/hou-2} }\newline
\subfloat[Set 3\label{fig:hou3}]{\includegraphics[width=\textwidth]{figures/hou-3} }\caption[Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right)]{Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).}\label{fig:hou}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
